---
title: "Lab 11: Cities"
author: "STAT 20 UC Berkeley"
output:
  xaringan::moon_reader:
    css: stat20.css
    lib_dir: assets/libs
    nature:
      highlightStyle: atelier-forest-light
      highlightLines: true
      highlightSpans: true
      ratio: 16:9
      countIncrementalSlides: false
      slideNumberFormat: "%current%"
---

```{r setup, include=FALSE, warning=FALSE}
#| include=FALSE,
#| warning=FALSE
knitr::opts_chunk$set(message = FALSE,
                      warning = FALSE,
                      echo = FALSE,
                      fig.align = "center",
                      fig.retina = 3)

library(tidyverse)
library(stat20data)
data(arbuthnot)
library(xaringanthemer)
source("https://raw.githubusercontent.com/stat20/course-materials/master/assets/stat20-style.r")
```

## Goals

- Describe the relation between two variables
- Give a more precise description of linear relations
- Interpret the relations and make predictions

---

## The data

We will use data about Brazilian cities. Our sample consists of a random sample of 60 cities from all 5573 cities in Brazil. It has been collected from various official websites and has been made available on kaggle.

We will try to predict the population of a city using a number of other features.

---
## The variables
```{r, echo = FALSE, eval = TRUE, message = FALSE, warning = FALSE}
cit <- read_csv("https://www.dropbox.com/s/vx3tmh3ybwtbqk7/cities.csv?dl=1")
dic <- read_csv("https://www.dropbox.com/s/pwbvn51x4o1fvh9/data_dic.csv?dl=1")
cit_all <- read_csv("https://www.dropbox.com/s/tbnac25c91krrhj/cities_all.csv?dl=1")
```

```{r, eval = TRUE, echo = FALSE}
dic %>% 
  select(Field, Description) %>%
  print()
```

---
# Part I

Q: What do you think the distribution of the variable `pop` looks like? Which plot would you use to show this distribution?

---
## Visualizing two numerical variables
Q: Choose a variable that you think has a strong relation to `pop`. How would you plot these two variables? Sketch a plot of what you expect. Describe how you expect the relation to look like. Is it linear? If yes, is it a positive or negative relation?

---
# Part II

## Loading the data

```{r, echo = TRUE, eval = TRUE}
library(tidyverse)
cit <- read_csv("https://tinyurl.com/4e57c5tj")
cit %>%
  glimpse()
```

---

##Visualizations
We now use a scatterplot to show the relation between the number of companies and the population of a city.

```{r, eval = TRUE, echo = FALSE}
cit %>% 
  ggplot(aes(x=comp, y = pop)) +
  geom_point() + 
  labs(y= "Population",
       x = "Number of companies")

```
How would you describe this relation?

---
We could see on the plot that there are a few outliers. More specifically there is a group of points that are close together and a few other ones that are well separated. Why could that be?

---

Let us now look at the distribution of the population size.
```{r, eval = TRUE, echo = FALSE}
library(patchwork)
p1 <- cit %>% 
  ggplot(aes(x=comp)) + 
  geom_histogram() + 
  labs(x = "Number of companies")
p2 <- cit %>% 
  ggplot(aes(x=pop)) + 
  geom_histogram() + 
  labs(x = "Population")
p1 + p2
```

How would you describe these distributions?

---
## Transforming variables

Because these variables are right-skewed we use the log-transform to make them more evenly spread out.
```{r, eval = TRUE, echo = FALSE}
p1 <- cit %>%
  ggplot(aes(x=log10(comp))) + 
  geom_histogram() + 
  labs(x = "log(Number of companies)")
p2 <- cit %>% 
  ggplot(aes(x=log10(pop))) +
  geom_histogram() + 
  labs(x = "log(Population)")
p1 + p2
```

---

We add these transformed variables to our dataframe:
```{r, eval = TRUE, echo = TRUE}
cit <- cit %>% 
  mutate(log_pop = log10(pop), 
         log_comp = log10(comp))
```

```{r, echo = F}
cit_all <- cit_all %>% 
  mutate(log_pop = log10(pop), 
         log_comp = log10(comp))
```


And look at the scatterplot of these variables:
```{r, eval = TRUE, echo = FALSE, fig.height = 4}
cit %>% 
  ggplot(aes(x=log_comp, y = log_pop)) + 
  geom_point() + 
  labs(y= "log(Population)",
       x = "log(Number of companies)")
```
Q: How would you describe this relation? What is its form, direction and strength? Is it linear?
---

## Correlation

Is used to assess the strength of a linear relation. 

For the correlation between the logarithm of the population and the logarithm of the number of companies we get:

```{r}
cit %>% 
  drop_na(log_pop, log_comp) %>% 
  summarize(cor(log_pop, log_comp))
```

---
## Correlation and visualization

Q: Use some other variable and compute its correlation with population. Is it stronger or weaker than you expected?

Q: Now plot the two variables using a scatterplot and describe their relation to each other. If the variables you selected are right-skewed, you might want to first create a new variable that is the logarithm of the new one.

---
## Manually fitting a line to a scatterplot

Just like we used the mean and standard deviation to summarize a single variable, we can summarize the relationship between these two variables by finding the line that best follows their association.

We will now use the interactive function `plot_ss` to manually fit a line to the points we have. Don't use this function in a markdown file!

We use this code:
```{r, eval = F, echo = T}
library(oilabs)
plot_ss(x = log_comp, y = log_pop, data = cit)
```
---

## Interpretting the fit of a line

The line you specified by clicking two points will be shown in
black and the residuals in blue. You will also be given an intercept, a slope and the sum of squared residuals. 

Recall that the residuals are the difference between the
observed values and the values predicted by the line:

$$
  e_i = y_i - \hat{y}_i
$$

The most common way to do linear regression is to select the line that minimizes the sum of squared residuals.

---
## Fitting a line to a scatterplot

We can use the `lm` function in R to fit the linear model that minimizes the sum of squared residuals.

```{r, eval = TRUE, echo = TRUE}
m1 <- lm(log_pop ~ log_comp, data = cit)
```

The first argument in the function `lm` is a formula that takes the form
`y ~ x`. Here it can be read that we want to make a linear model of `log_pop` as a function of `log_comp`. The second argument specifies that R should look in the `cit` data frame to find the `comp` and `pop` variables.

---
## The coefficients of a linear model

The output of `lm` is an object that contains all of the information we need about the linear model that was just fit. Let us look at the coefficients.

```{r, eval = TRUE, echo = TRUE}
coefficients(m1)
```

So the line with the best fit is approximately
$$ \log(pop) = 2.54 + 0.69 \cdot \log(comp).$$
---
## Computing the slope

We can also compute the slope from two stats we have already seen: the SD of the two variables and their $r$.

```{r}
cit %>%
  summarize(sd_p = sd(log_pop), 
            sd_c = sd(log_comp), 
            cor_pc = cor(log_pop,log_comp)) %>%
  mutate(slope = sd_p/sd_c*cor_pc)
```
Q: Find the slope of the linear regression line that describes the relation between the variable you chose earlier and `pop` using this method.

---
## Predictions

Let us see how well this model predicts the population of cities that were not in the dataset.

```{r, echo = F}
set.seed(4)
cit_new <- cit_all %>% 
  filter(!(city %in% cit$city)) %>% 
  select(city, log_comp, log_pop) %>%
  slice_sample(n=3)
cit_new %>% 
  select(-log_pop)
```
Q: What is the predicted population size for these cities according to our model? 

---
Instead of doing this by hand we can also use the `predict` function in R. It takes the model as the first input and the new rows as the second input.
```{r, echo = T, eval = T}
predict(m1, cit_new)
```
Let us compare these values to the true ones:

```{r, echo = FALSE}
cit_new %>% 
  mutate(log_pop_pred = predict(m1,cit_new))
```
