---
title: "Slide Template"
author: "STAT 20 UC Berkeley"
output:
  xaringan::moon_reader:
    css: stat20.css
    lib_dir: assets/libs
    seal: false
    nature:
      highlightStyle: atelier-forest-light
      highlightLines: true
      highlightSpans: true
      ratio: 16:9
      countIncrementalSlides: false
      slideNumberFormat: "%current%"
---

```{r setup, include=FALSE, warning=FALSE}
knitr::opts_chunk$set(message = FALSE,
                      warning = FALSE,
                      echo = FALSE,
                      fig.align = "center",
                      fig.retina = 3)

library(tidyverse)
library(xaringanthemer)
source("https://raw.githubusercontent.com/stat20/course-materials/master/assets/stat20-style.r")
```

class: center, middle

# Welcome to Stat 20!
### Introduction to Probability and Statistics

Lab 2

---
## Components of RStudio

1. Console
--


2. Environment
--


3. Editor
--

4. File Directory
--

.task[
Now we are going to switch over to datahub to understand these 4 components a bit better
]

---
## R as a calculator 

The basic mathemetical operators `+`, `-`, `*`, `/`, `^` can be used directly in the R console or a .R script or .RMD
--

```{r addition, echo = TRUE, eval = FALSE}
4 + 3
```

--

```{r ref.label = "addition", echo = FALSE}
```

--

```{r exp, echo = TRUE, eval = FALSE}
2 ^ 3
```

--

```{r ref.label = "exp", echo = FALSE}
```

---
## Assignment and Help Operators

The assignment operator `<-` is a way to assign variables in R. It is the equivalent of `=` in other programming languages. 
--

```{r hello, echo = TRUE, eval = FALSE}
var <- "hello"
```

--

```{r ref.label = "hello", echo = FALSE}
```

--

The help operator is a convenient way to get more information in RStudio on a function of interest

1. `?mean`
2. `help(mean)`

---

## Vectors and Tibbles (Data Frames)

.pull-left[
Vectors is the simplest data element used in R to store an individual variable. One way they can be created is using the `concatenate` operator:


```{r cat, echo = TRUE, eval = FALSE}
vector <- c("a","b","C")
print(vector)
```

```{r ref.label = "cat", echo = FALSE}
```
]

--

.pull-right[
Tibbles / data frames are a collection of vectors in R that can contain multiple data types.

`class()` can be used to see what data types are presents


```{r class, echo = TRUE, eval = FALSE}
vector <- c("a","b","C")
class(vector)
```

```{r class, echo = FALSE, eval = TRUE}
```

]

---

## Tibbles (Data Frames) Continued

Ways to load pre-existing data frames into R is using 

1. `data()`
2. `tibble()`

```{r tibble, echo = TRUE, eval = FALSE}
tibble(mtcars)
```

--

```{r tibble2, echo = FALSE, eval = TRUE}
head(tibble(mtcars))
```

---

## Basic Functions on Vectors / Tibbles

Functions can be called on vectors and columns in a tibble. For example `sum()` and `mean()`

```{r math, echo = TRUE, eval = FALSE}
vector <- c(1,2,3,4,5)
sum(vector)
mean(vector)
```

--

```{r ref.label = "math", echo = FALSE}
```

---

## Packages in R

There are several additional "packages" in R that can be used to make frequently done tasks much easier. One such package is the `dplyr` package in `tidyverse`. 

Packages can be loaded into R using `library()`

```{r lib, echo = TRUE, eval = FALSE}
library(tidyverse)
```

---

## Data Manipulation using `tidyverse`

There are several functions that can be used to manipulate tibbles easily: 

* `select()` : subset columns 

--


* `arrange()` : sort columns 

--


* `mutate()` : create new columns from existing columns

--

---

## Data Manipulation Example

Using the functions that we have introduced, take some time to complete the following: 

.task[Please provide the R code that will generate a tibble with the following characteristics: - Three variables: Name, Favorite Color, Favorite Number - 5 observations.]

The code should save this tibble in an object called `my_classmates`.
