---
title: "Lab 1"
author: "Andrew Bray"
date: '2022-02-02'
output: pdf_document
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE, fig.align = "center", fig.width = 4, fig.height = 2.5)
```

## My Home: San Diego Country, CA

San Diego County, officially the County of San Diego, is a county in the southwestern corner of the state of California, in the United States. As of the 2020 census, the population was 3,298,634, making it California's second-most populous county and the fifth-most populous in the United States. Its county seat is San Diego, the second-most populous city in California and the eighth-most populous city in the United States. It is the southwesternmost county in the 48 contiguous United States, and is a border county.

## My favorite spots

1. Moonlight Beach
2. Juanita's
3. Pannikin Coffee and Tea

## My favorite animals

I really like penguins. Let me show you some data visualization about their beak sizes.

```{r}
library(tidyverse)
library(palmerpenguins)
data(penguins)
ggplot(penguins, aes(x = bill_length_mm,
                     y = bill_depth_mm,
                     color = island)) +
  geom_point()
```

As you can see above, there is a generally positive association between bill length and bill depth in penguins found on each of the islands. Penguins from Biscoe have shorter bills than penguins on other islands.


