---
title: "Lab 1: Getting Started"
author: "STAT 20 UC Berkeley, Spring '22"
output:
  xaringan::moon_reader:
    lib_dir: "assets/libs"
    css:
      - assets/css/stat-375-slides.css
    nature:
      slideNumberFormat: "%current%"
      highlightStyle: solarized-light
      highlightLines: true
      ratio: 16:9
      countIncrementalSlides: true
---

```{r setup, include=FALSE, warning=FALSE}
knitr::opts_chunk$set(message = FALSE, 
                      fig.align = "center",
                      fig.retina = 3,
                      warning = FALSE)

library(tidyverse)
library(emo)
library(xaringanExtra)
library(countdown)
xaringanExtra::use_panelset()
```

## Agenda 

- Icebreaker + Introduction (10 min) 
- Intro to Ed (5 min)
- Lab 1 Demo (10 min)
- Lab 1 Students (20 min)
- Q & A (5 min)

---
## Icebreaker
--

1. Write down 2 truths and a lie (1 min)

--

2. In breakout rooms of size 4, take turns sharing:
  - name
  - hometown
  - 2 truths and lie
  
--

3. After each student shares, group tries to guess which is the lie

```{r echo = FALSE}
countdown::countdown(1)
```

---
## What is Stat 20 Lab?

.pull-left[
**Who**: 25 students + GSI

**What**: applying the concepts from lecture by working with real data  

**When**: Tu, Th mornings for 50 minutes  

**Where**: Initially Zoom, then 3rd floor Evans  
]

--

.pull-right[
**Why**: Doing statistics requires a synthesis of conceptual frameworks, ideas and notation from mathematics, computation tools, and an understanding of the context of the data. It's a complex task and it requires practice.  

**How**: When doing data analysis, for our computational tool we'll be using R.
]

---
