---
title: "Lab 4: Flights II"
author: "STAT 20 UC Berkeley"
output:
  xaringan::moon_reader:
    css: stat20.css
    lib_dir: assets/libs
    seal: false
    nature:
      highlightStyle: atelier-forest-light
      highlightLines: true
      highlightSpans: true
      ratio: 16:9
      countIncrementalSlides: false
      slideNumberFormat: "%current%"
---

```{r setup}
#| include=FALSE,
#| warning=FALSE
knitr::opts_chunk$set(message = FALSE,
                      warning = FALSE,
                      echo = FALSE,
                      fig.align = "center",
                      fig.retina = 3)

library(tidyverse)
library(xaringanthemer)
source("https://raw.githubusercontent.com/stat20/course-materials/master/assets/stat20-style.r")
```

class: center, middle

# Lab 4: Flights II
### Introduction to Probability and Statistics


---
## Agenda

--

1. Review group_by()  and summarize()

--

2. Flights data with ggplot

---


class: middle

# group_by() and summarize()

---

## Flights by month

Describe the steps to create the following summary from the flights data. Specify the functions and variables.

--

```{r}
library(stat20data)
data(flights)
flights %>%
  group_by(month) %>%
  summarise('Number of Flights' = n())
```

---

class: middle

# Part II: Computing with the Data - continued

---
## Summarize the data

16. Create a data frame that contains the median and interquartile range for departure delays, grouped by carrier. Which carrier has the lowest typical departure delay? Which one has the least variable departure delays?

---
## Simple ggplots

7. Create a bar chart that shows the distribution by month of all the flights leaving the Bay Area (SFO and OAK). Do you any sign of an effect of the pandemic?

---
## dplyr + ggplot 

12. Create a plot that captures the relationship of average speed vs. distance and describe the shape and structure that you see. What phenomena related to taking flights from the Bay Area might explain this structure?

9. Add a new column to your data frame called before_times that takes values of TRUE and FALSE indicating whether the flight took place up through the end of March or after April 1st, respectively. Remake the histograms above, but now separated into two subplots: one with the departure delays from the before times, the other with the flights from afterwards. Can you visually detect any difference in the distribution of departure delays?



