---
title: "PS 3"
output:
  pdf_document: default
  html_document:
    df_print: paged
---


```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = FALSE,
                      message = FALSE)
```

1. **Raise taxes.** 
A random sample of registered voters nationally were asked whether they think it's better to raise taxes on the rich or raise taxes on the poor. 
The survey also collected information on the political party affiliation of the respondents. [@survey:raiseTaxes:2015]

    ```{r fig.asp = 0.4}
    library(tidyverse)
    library(scales)
    library(openintro)
    
    raise_taxes_survey <- tribble(
       ~"Opinion",                ~"Democrat", ~"Republican", ~"Independent/Other",
       "Raise taxes on the rich", 91,	       47,            49,
       "Raise taxes on the poor",  4,	       10,            11,
       "Not sure",                 5,	       43,            40
    )
    
    raise_taxes_survey %>%
      pivot_longer(
            cols = -Opinion,
            names_to = "Party",
            values_to = "Proportion"
      ) %>%
      mutate(Opinion = fct_relevel(Opinion, 
                                   "Raise taxes on the rich", 
                                   "Raise taxes on the poor", 
                                   "Not sure")) %>%
      ggplot(aes(y = Party, x = Proportion, fill = Opinion)) +
      geom_col() +
      scale_fill_manual(values = c(IMSCOL["blue", "full"], 
                                   IMSCOL["red", "full"], 
                                   IMSCOL["lgray", "full"])) +
      scale_x_continuous(labels = label_percent(scale = 1)) +
      labs(x = NULL) +
      theme(legend.position = "bottom") +
      guides(fill = guide_legend(reverse = TRUE))
    ```

    a. Based on the stacked bar plot shown above, do views on raising taxes and political affiliation appear to be associated? Explain your reasoning.

    b. Conjecture other possible variables that might explain the potential association between these two variables.


2. **Associations.**
Indicate which of the plots show (a) a positive association, (b) a negative association, or (c) no association. Also determine if the positive and negative associations are linear or nonlinear. Each part may refer to more than one plot.

    ```{r fig.asp=0.5}
    library(tidyverse)
    set.seed(2306)
    
    # create x
    x <- seq(0, 10, 0.1)
    
    # create y_poslin: positive linear with x
    y_poslin <- x * runif(1, min = 0, max = 10) + 
      rnorm(length(x), mean = 0, sd = runif(1, min = 3, max = 10)) - 
      runif(1, min = 0, max = 3)
    
    # create y_neglin: negative linear with x
    y_neglin <- x * -runif(1, min = 0, max = 5) + 
      rnorm(length(x), mean = 0, sd = runif(1, min = 3, max = 10)) - 
      runif(1, min = 0, max = 5)
    
    # create y_poscur: curved positive with x
    y_poscur <- x^2 + rnorm(length(x), mean = -3, sd = runif(1, min = 3, max = 4))
    
    # create y_none: no association with x
    y_none <- x + rnorm(length(x), mean = 0, sd = runif(1, min = 30, max = 40))
    
    # data frame
    tibble(
      x = x,
      poslin = y_poslin,
      neglin = y_neglin,
      poscur = y_poscur,
      none   = y_none
    ) %>%
      pivot_longer(cols = -x, names_to = "type", values_to = "y") %>%
      arrange(type) %>%
      mutate(
        type_num = case_when(
          type == "poslin" ~ "(1)",
          type == "neglin" ~ "(4)",
          type == "poscur" ~ "(3)",
          type == "none"   ~ "(2)"
        ),
        type_num = as.factor(type_num)
        ) %>%
      ggplot(aes(x = x, y = y)) +
      geom_point(alpha = 0.5) +
      facet_grid(~type_num) +
      theme(
        axis.text = element_blank()
      ) +
      labs(x = NULL, y = NULL)
    ```
  
\newpage  
    
3. **Infant mortality.** 

The infant mortality rate is defined as the number of infant deaths per 1,000 live births. This rate is often used as an indicator of the level of health in a country. 
The relative frequency histogram below shows the distribution of estimated infant death rates for 224 countries for which such data were available in 2014.

```{r}
    library(ggplot2)
    library(scales)
    library(openintro)
    
    ggplot(subset(cia_factbook, !is.na(cia_factbook$infant_mortality_rate)), aes(x = infant_mortality_rate)) +
      geom_histogram(aes(y = stat(count) / sum(count)),
        binwidth = 10
      ) +
      labs(
        x = "Infant mortality (per 1,000 live births)", 
        y = "Proportion") + theme_minimal()
```

    a.  Estimate Q1, the median, and Q3 from the histogram.

    b.  Would you expect the mean of this dataset to be smaller or larger than the median? Explain your reasoning.

\newpage
  
4. **Median vs. mean.** 
Estimate the median for the 400 observations shown in the histogram and note whether you expect the mean to be higher or lower than the median.

    ```{r}
    library(tidyverse)
    library(openintro)
    
    set.seed(9823)
    df <- tibble(x = (100 * rbeta(400, 12, 3)))
    
    ggplot(df, aes(x = x)) +
      geom_histogram(binwidth = 5) +
      labs(x = "", y = "") +
      scale_y_continuous(breaks = seq(0, 80, 20)) + theme_minimal()
    ```


5. **R code** 

Using ggplot in R code: plot the histogram of mpg of mtcars, and detemine whether the median is larger, smaller, or equal to the mean. (Show your code)