---
title: "Lab 7B: Hypothesis Testing"
author: "STAT 20 UC Berkeley"
output:
  xaringan::moon_reader:
    css: stat20.css
    lib_dir: assets/libs
    seal: false
    nature:
      highlightStyle: atelier-forest-light
      highlightLines: true
      highlightSpans: true
      ratio: 16:9
      countIncrementalSlides: false
      slideNumberFormat: "%current%"
---

```{r setup, include=FALSE, warning=FALSE}
#| include=FALSE,
#| warning=FALSE
knitr::opts_chunk$set(message = FALSE,
                      warning = FALSE,
                      echo = TRUE,
                      fig.align = "center",
                      fig.retina = 3)

library(tidyverse)
library(stat20data)
data(class_survey)
library(xaringanthemer)
source("https://raw.githubusercontent.com/stat20/course-materials/master/assets/stat20-style.r")
```

class: center, middle

# Lab 8: Elections
### Introduction to Probability and Statistics

???

## Timing

[15 min] Introduction to Benford's Law using primarily boardwork
[3 min] Introduction to Iran Election context (slides)
[15 min] Students work on EDA portion
[5 min] Review as a class things that came up during EDA
[5 min] Show how to use first digit function (slides)
[remaining] Students continue working on remaining problems

This first component can be done either during Berkeley time or at the start of class along with students. It depends on how long you expect the rest of the lesson to take and what you'd like to emphasize:
[Write on the board the two "fundamental" distributions that we've covered so far: the Binomial and the Bernoulli. Slides 30 and 55 of fundamental_distributions.html will give you a sense of the notation and layout that's used. Best to replicate each element of these slides, including the plots.] 

It's perfectly reasonable to ask students to close their laptops for the first part of class.

Ask students for a guess or two at the population of Berkeley.  Write on board then go to Wikipedia and get the true answer and correct the one(s) on the board. From there, follows links through Wikipedia to other towns and cities in California, building up a list of 10 city names and their populations (randomly picking a few of these links is a good method: https://en.wikipedia.org/wiki/Category:Incorporated_cities_and_towns_in_California). Structure this on the board as a data frame with two columns: city and population, and 10 rows.

Ask students to sketch the distribution of two variables:
1. What they expect the population variable to look like were they to collect the population of all the cities and towns in California.
2. The distribution of the value of the first digit in the population counts of all of these town.
This is probably best done think-pair share: give a couple minutes to for them to sketch silently, then ask them to share with a neighbor, then ask a pair to describe their distributions as you draw it on the board.

Probing questions:
1. Where does this variable sit in the Data Taxonomy?
2. What is the range of possible values this variable can take?
3. What geometry will you use?
4. What shape/modality/center/spread would you expect to see?
5. What labels belong on the axes?
6. What is a good title for this plot?

The first should be a histogram or density plot (boxplot is meh) that is heavily right skewed. The second should be a barchart on the integers 1-9, each one having decreasing probabilities.

Write out Benford's Law in a similar way to the other named distributions and describe it as a distribution that we're going to try to use to describe vote counts. Start the calculation of E and V of Benford's Law just so that students can see what goes into each term of those sums (I'm actually not sure what the base_10 Benfords E and V are).

---

# Agenda

1. Benford's Law
2. 2009 Iran Election
3. Exploring election data
5. Benford's Law in Iran


---
class: middle, center, inverse

## Benford's Law

Observation: many naturally occurring numerical variables have a recurring pattern in the distribution of the first digit.

Benford's law observes that in many naturally-occuring phenomena with a quantitative measurement, some patterns can be observed in the digits of that measurement. Specifically, the law states that the first digit of such observations, rather than being distributed uniformly between the numbers 1-9, takes a decreasing log distribution. The code and plotted distribution for this are below. Investments with interest, populations of cities, and election results have all been observed by different groups to follow Benford's law.

```{r benford_code, echo = TRUE, eval = TRUE}
benfords_p <- data.frame(first_digit = 1:9, 
                         ben_prop = log10(1 + 1/1:9))
```

```{r benford_plot, eval = TRUE}
library(tidyverse)
colnames(benfords_p) <- c("digit", "prop")
ggplot(benfords_p, aes(x = digit, y = prop)) +
  geom_bar(stat = "identity") +
  scale_x_discrete(limits = 1:9)
```

---
class: middle, center, inverse

# 2009 Iran Election

---
class: middle, center, inverse

# Exploring election data

---
## Exploring election data

```{r}
countdown::countdown(15)
```

---
class: middle, center, inverse

# Benford's Law in Iran

---
## First Digits

```{r hint, echo = TRUE, eval = FALSE}
substr("A long time ago, in a galaxy far far away...", 21, 28)
substr(387, 1, 2)
as.numeric(substr(387, 1, 2))
get_first(387)
```

---
## Distribution of first digits in Iran

```{r eval = TRUE, echo = FALSE, fig.width=12, fig.height=6, fig.align='center'}
library(stat20data)
library(patchwork)
data(iran)
p1 <- iran %>%
  mutate(first_digit = get_first(ahmadinejad)) %>%
  ggplot(aes(x = first_digit)) +
  geom_bar() +
  labs(x = "first digit",
       title = "Ahmadinejad Vote Data")

benfords_p <- data.frame(first_digit = 1:9, 
                         ben_prop = log10(1 + 1/1:9))
colnames(benfords_p) <- c("digit", "prop")
p2 <- ggplot(benfords_p, aes(x = digit, y = prop)) +
  geom_bar(stat = "identity") +
  scale_x_discrete(limits = 1:9) +
  labs(title = "Benford's Law")
p1 + p2
```


